---
title: "Time Series Analysis Assignment 11"
author: "Luay Dajani"
date: "April 7, 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## Libraries

# install.packages("dygraphs")
if("dygraphs" %in% rownames(installed.packages()) == FALSE) {install.packages("dygraphs")}
library("dygraphs")

# install.packages("forecast")
if("forecast" %in% rownames(installed.packages()) == FALSE) {install.packages("forecast")}
library("forecast")

# install.packages("fpp2")
if("fpp2" %in% rownames(installed.packages()) == FALSE) {install.packages("fpp2")}
library("fpp2")

# install.packages("xts")
if("xts" %in% rownames(installed.packages()) == FALSE) {install.packages("xts")}
library("xts")
```

# Summary

The following shows fun with Time Series and Prediction using exponention smoothing forecasts (SES and HW), ploting with regular plots as well as dygraphs

GITHub Repository: https://github.com/LuayD/SMU-MSDS/tree/master/TimeSeries_6306_11

## Analysis 1: Warm-up Bried Financial Data from EU Stock Markets

This part takes the built in Eu Stock Markets Data Set and isolates the DAX listings, plotting observed and decomposing using decompose to show seasonal, trend, and random trends of the data.

```{r analysis1}
## Part 1a. Pull data from EU Stock Markets 1991 - 1998

# Take just the "DAX" data from the data
ts_EuStockMarketsDax <- EuStockMarkets[,1]
names(ts_EuStockMarketsDax) <- "DAX"

## Part 1b. Plot with bluw line and make red vert line on 1997
plot(ts_EuStockMarketsDax,
     main = "EU Stock Market DAX index 1991 - 1998",
     ylab = "DAX index",
     xlab = "Year",
     col = "blue")
abline(v=1997, col="red", lwd = 2)

## Part 1c. decompose data "mult" to get observed, seasonal, trend, & random noise, same colors as before
lst_decomp_EuStockMarketsDax <- decompose(ts_EuStockMarketsDax, type="mult")
plot(lst_decomp_EuStockMarketsDax, 
    xlab = "Year",
    col = "blue",
    las = 1,         # make Categorylabel text perpendicular to axis & plot
    mar = c(0,9,1,3) # increase y-axis margin (c(bottom, left, top, right))
)
abline(v=1997, col="red", lwd = 2)
```

## Analysis 2: Temperature Data Predicitons for Moorabbin Airport Melbourne

This part uses Exponential smoothing forecasts SES and HW to predict Airport temperatures for subsequent 5 years and plots.

```{r analysis2}
## Part 2a. Pull data from fpp2 lib and plot

# assign to local var to visually explore
ts_maxtemp <- maxtemp

# Autoplot maxtemp data using example as per help page
autoplot(ts_maxtemp, main = "Max Temperate for Moorabbin Airport Melbourne",
         xlab = "Time Period", ylab = "Temperature (Celcius")

## Part 2b. subset data after year 1990
ts_maxtemp_1990 <- window(ts_maxtemp, start = 1990)

## Part 2c. Predict next 5 years of temperatures

#Get predicted meand part 5 years
lst_maxtemp_ses_1990_2021 <- ses(ts_maxtemp_1990, h = 5)

# plot observed & 5 years predicted with confidence intervals of 80% and 95%
plot(lst_maxtemp_ses_1990_2021, 
     main = "Moorabbin Airport 1990-2016 +5 years (SES) forecasted",
     sub = "Black Observed & Blue Fitted (1990-2016) Predicted (2017-2021 w/ 80% & 95% CI)",
     ylab = "Temperature (Celcius)",
     xlab = "Year",
     type = "o"
     )

# find fitted (predicted values lines) from 1990 - 2016 (NOT THE 5 YEARS PREDICTED) & add line to plot
lines(fitted(lst_maxtemp_ses_1990_2021), col = "blue", type = "o")

# overlay 2016-2021 predictions with points
lines(lst_maxtemp_ses_1990_2021$mean, col = "blue", type = 'o')

## Part 2d. Use Damped Holt's Linear Trend
lst_maxtemp_holt_1990_2021 <- holt(ts_maxtemp_1990, h = 5, damped = TRUE)

# plot observed & 5 years predicted with confidence intervals of 80% and 95%
plot(lst_maxtemp_holt_1990_2021, 
     main = "Moorabbin Airport 1990-2016 +5 years (Holt) forecasted",
     sub = "Black Observed & Blue Fitted (1990-2016) Predicted (2017-2021 w/ 80% & 95% CI)",
     ylab = "Temperature (Celcius)",
     xlab = "Year"
)

# find fitted (predicted values lines) from 1990 - 2016 (NOT THE 5 YEARS PREDICTED) & add line to plot
lines(fitted(lst_maxtemp_holt_1990_2021), col = "blue", type = 'o')

# overlay 2016-2021 predictions with points
lines(lst_maxtemp_holt_1990_2021$mean, col = "blue", type = 'o')

# AICc comparion for models
cat("AICs for SES model", lst_maxtemp_ses_1990_2021$model$aicc, "\n")
cat("AICs for HW model", lst_maxtemp_holt_1990_2021$model$aicc, "\n")
cat("The SES model has the lower AICc", "\n")
```

## Analysis 3: Wands that choose the wizard

This part reads in Harry Potter based Wand makers and shows output per year using an interactive dygraph with output show in the period where "He who should not be named" was active.

```{r analysis3}
## Part 3a.Read in both datasets, files do not have headers
dfm_ollivander <- read.csv("data/Unit11TimeSeries_Ollivander.csv", header = FALSE, stringsAsFactors = FALSE)
dfm_gregorovitch <- read.csv("data/Unit11TimeSeries_gregorovitch.csv", header = FALSE, stringsAsFactors = FALSE)

# Add headers
colnames(dfm_ollivander) <- c("year", "ollivander")
colnames(dfm_gregorovitch) <- c("year", "gregorovitch")

## Part 3b. Convert aquired Date into correct date format
dfm_ollivander["newdate"] = as.Date(dfm_ollivander$year, "%m/%d/%Y")
dfm_gregorovitch["newdate"] = as.Date(dfm_ollivander$year, "%m/%d/%Y")

# Combine (Merge) data frames
dfm_combined <- merge(dfm_ollivander[c("newdate","ollivander")], 
                                     dfm_gregorovitch[c("newdate","gregorovitch")], 
                                     by = "newdate", all=TRUE)

## Part 3c. Convert data frames to time series and plot using dygraphs
ts_combined <- xts(dfm_combined[c("ollivander","gregorovitch")], 
                   order.by = dfm_gregorovitch$newdate)


# plot the dygrpah
dygraph(ts_combined, main="Wands Sold per Wand Maker", ylab="Wands Sold", 
        xlab="Year (Shaded Area When He who should not be named active)") %>% 
  dySeries("ollivander", label = "Ollivander", color = "green") %>%
  dySeries("gregorovitch", label = "Gregorovitch", color = "red") %>%
  dyOptions(stackedGraph = TRUE) %>%
  dyRangeSelector(height=30) %>%
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>%
  dyShading(from = "1995-1-1", to = "1999-1-1", color = "#55CCDD")
```

## References from comments in the code

##### 1- dygraphs for R: https://rstudio.github.io/dygraphs/index.html
##### 2- Check for installed packages before running install.packages(), 2 down: https://stackoverflow.com/questions/9341635/check-for-installed-packages-before-running-install-packages

## Environment Notes

```{r environment_notes, echo=FALSE}
sessionInfo()
```