---
title: "LuayDajani-MSDS6306-HW4"
author: "Luay Dajani"
date: "February 6, 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 6306 Assignment 4 for Luay Dajani

Homework done as per specifications with Markdown formatted as an HTML file.

This work available on the Github repository: https://github.com/LuayD/SMU-MSDS/tree/master/SMU-MSDS-6306-HW4

Please note data for this homework from fivethirtyeight available at github: https://github.com/rudeboybert/fivethirtyeight 

## Questions 1: FiveThirtyEight Data

### 1a: Setup Environment and install packages

```{r 1a, echo = FALSE}
#Print out R-Package information
print(paste("Using", R.version.string))
#Package Installs
install.packages("fivethirtyeight", dependencies=TRUE, repos='http://cran.rstudio.com/')
library(fivethirtyeight)
```

### 1b: Load 18th dataset in library

```{r 1b}
#set 18th data set into "df"
df <- college_recent_grads
```

### 1c: Display detailed list of data sets for the package

```{r 1c}
#use vingette and print out URL
#vin <- browseVignettes(package = "fivethirtyeight") #could not figure this out!!! so went ahead and out the URL of interest from the packsge below:

print(c("Please see more information about this topic at  http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major"))
```

### 1d: Give dimentions & Column names of data frame

```{r 1d}
#print dimentions
print(paste("The data set college_recent_grads in fivethirtyeight has"
            , dim(df)[2]
            , "Columns, and"
            , dim(df)[1]
            , "Rows."
))
#print column names
print("The column names include the following:")
cat(names(df), sep = ", ")
```

## Questions 2: FiveThirtyEight Data Summary

### 2a: Write Column names and Column Number

```{r 2a}
#print column names and Number
cat("Again, the column names include the following:\n"
    , names(df)
    , "\nThe numbers of column in this data set are: "  
    , length(names(df))
)
```

### 2b: give the unique listing for major categories in the df and frequency they show up

```{r 2b}
#print column names and Number
library(plyr)
major_count <- count(df, 'major_category')
major_count
```

### 2c: Barplot of unique major categories and frequency

```{r 2c}
#create the barplot with label, title, horizontal
par(las = 2)
par(mar=c(5,10,4,2)) # increase y-axis margin (c(bottom, left, top, right))
barplot(major_count$freq
        , main="Recent College Grads & Categories for Majors"
        , horiz=TRUE
        , xlab="Frequency in each category"
        , cex.names=0.6
        , names.arg = major_count$major_category
        , col=c("red","green","blue","orange","brown","purple","yellow")
        )
```

### 2d: fivethirtyeight recent college grads -> csv file w/o row labels

```{r 2d}
#write to csv file df "recent college grads" data
write.csv(df
          , "fivethirtyeight_recent-college-grads.csv"
          , row.names = FALSE
          )
```

## Question 3: Codebook

### 3a: Add New Repository on Github
Available at: https://github.com/LuayD/SMU-MSDS

### 3b: Add README.md to root 
Done and Available at: https://github.com/LuayD/SMU-MSDS

### 3c: Add nested README to and assignment folder and credit fivethirtyeight
Done and Available at: https://github.com/LuayD/SMU-MSDS/tree/master/SMU-MSDS-6306-HW4

### 3d: Link to Github:
Done, and mentioned multiple times in this document. Relevant folders include:
-Main root section: https://github.com/LuayD
-Root SMU MSDS repository section: https://github.com/LuayD/SMU-MSDS 
